---
title: "Advanced Programming with R - Homework 1"
author: "Pranav Agwan, Student No - 24219261"
format: pdf
editor: visual
---

# [***Titanic Disaster***]{.underline}

In this section we load the **Titanic** dataset which contains the **survival data** of passengers from the **Titanic disaster**. It **summarizes** the counts of **passengers** based on **class**, **age**, **sex** and **survival status**.

```{r}
#| echo: false
# Loading the dataset
data("Titanic")

# Converting the dataset to a dataframe
data = as.data.frame(Titanic)

# Displaying the top 6 values of the dataset
head(data)
```

We can see the top 6 values of our data above. Now we will see the structure.

```{r}
#| echo: false
# Checking the structure of the dataset
str(data)
```

There are total **32 observations** of **5 variables** in the **Titanic dataset.** The variables are as follows:

-   [***Class***]{.underline} - It shows the **passenger class**. It is a **factor variable** with 4 levels: **"1st"**, **"2nd"**, **"3rd"** and **"Crew"**.

-   [***Sex***]{.underline} - It shows **passenger's sex**. It is also a **factor variable** with 2 levels: **"Male"** and **"Female"**.

-   [***Age***]{.underline} - It shows the **age group** of the **passenger**. It is also a **factor variable** with 2 levels: **"Child"** and **"Adult"**.

-   [***Survived***]{.underline} - It shows the **survival status** of the **passenger**. It is also a **factor variable** with 2 levels: **"Yes"** and **"No"**.

-   [***Freq***]{.underline} - It shows the **count** of the **passengers** that fall into the specific groups mentioned above. It is a **numeric variable**.

We will now show a simple **bar plot** which displays the **numbe**r of **male** and **female** passengers who **survived** or **did** **not survived** the incident. We will use the **Sex**, **Survived** and **Freq** column for the plot.

```{r}
#| echo: false
# Loading the required library
library(ggplot2)

# Aggregating the number of survivals based on sex
survival = aggregate(Freq ~ Sex + Survived, data = data, sum)

# Plotting the graph
ggplot(survival, aes(x = Sex, y = Freq, fill = Survived)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Titanic Survival by Sex",
       x = "Sex",
       y = "Passenger Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

The **red bar** shows the passenger which **did not survived** the incident and the **blue bar** shows the **passenger** which survived. There are **2 bars** both for "**Male"** and **"Female"**. From the above graph we can interpret the following:

-   There are aprrox **1350-1400 males** which **did not survived** the incident.

-   There are approx **350-400 males** which **survived** the incident.

-   There are approx **100-150 females** which **did not survived** the incident.

-   There are approx **350-400 females** which **survived** the incident.
